Topological Graph Theory

### 2019-05-10 Topology/Paper/paper.tex
LETS DO THIS!

%\subsection{Neighborhood}
%Given some graph $G$ with vertex $v$. The neighborhood of $v$ within $G$ is the set of all vertices that share an edge with $v$, not including $v$.\footnote{If $G$ is a multigraph the definition stays the same, however $v$ maybe it's own neighbor if it has an edge with itself or a loop} Given a subgraph $H$ of $G$, the neighborhood of $H$ within $G$ is a new subgraph of $G$ that includes $H$ as well as any vertex in $G$ that is adjacent to a vertex in $H$ and the edge that they share.
%
%Lets look at some examples
%\begin{center}
%	\begin{tikzpicture}[auto, node distance=3cm, every loop/.style={},
%	thick,main node/.style={circle,draw,font=\sffamily\Large\bfseries}]
%	
%		\node[main node] (0) {\textcolor{blue}{$a$}};
%		\node[main node] (1) [above right of=0] {\textcolor{red}{$b$}};
%		\node[main node] (2) [right of=1] {$c$};
%		\node[main node] (3) [below right of=2] {$d$};
%		\node[main node] (4) [below of=3] {\textcolor{red}{$e$}};
%		\node[main node] (5) [below left of=4] {$f$};
%		\node[main node] (6) [left of=5] {$g$};
%		\node[main node] (7) [above left of=6] {$h$};
%		
%		\path[every node/.style={font=\sffamily\small}]
%		(0)
%		edge node {} (2)
%		edge node {} (3)
%		edge node {} (4)
%		(1)
%		edge node {} (3)
%		edge node {} (4)
%		edge node {} (5)
%		(2)
%		edge node {} (4)
%		edge node {} (5)
%		(3)
%		edge node {} (7)
%		(4)
%		edge node {} (6)
%		edge node {} (7)
%		(5)
%		edge node {} (6)
%		edge node {} (7)
%		(6)
%		edge node {} (7);
%	\end{tikzpicture}
%\end{center}
%
%First let us consider the neighborhood of just the vertex \textcolor{blue}{$a$}, this would then be all the vertices that it is adjacent to so we get $$\set{c, d, \textcolor{red}{e}}$$ We could also let $A$ be a subgraph containing only the vertex $\textcolor{blue}{a}$ and so it's neighborhood within $G$ will be the graph below.
%\begin{center}
%	\begin{tikzpicture}[auto, node distance=3cm, every loop/.style={},
%	thick,main node/.style={circle,draw,font=\sffamily\Large\bfseries}]
%	
%		\node[main node] (0) {\textcolor{blue}{$a$}};
%		\node[main node] (2) [above right of=0] {$c$};
%		\node[main node] (3) [below right of=2] {$d$};
%		\node[main node] (4) [below of=3] {\textcolor{red}{$e$}};
%		
%		\path[every node/.style={font=\sffamily\small}]
%		(0)
%		edge node {} (2)
%		edge node {} (3)
%		edge node {} (4)
%		(2)
%		edge node {} (4);
%	\end{tikzpicture}
%\end{center}
%The difference is that this is actually a subgraph of $G$, whereas the previous is just vertices.
%
%For another example let $H$ be the subgraph with only the \textcolor{red}{$b$}, \textcolor{red}{$e$} and the edge they share. The neighborhood of $H$ within $G$ within $G$ would then be all of $G$ as every vertex is either a neighbor of $b$ or $e$.
%
\subsection{Contraction}
This is not a property, but rather an operation or action that we preform on a graph. Let us consider the following graph.

@@

We wish to preform a contraction on edge $R$. To be clear all graph contractions are on edges. So we will make a new vertex, $r\cdot s$ which has all the edges of $r$ and all the edges of $s$ except the edge we are contracting across. In this case the edge we are contracting across is $R$ and thus we get the following multigraph.

@@

This can then be reduced into a graph again by simply treating $b_r$ and $b_s$ as the same edge. This operation is particularly important as we will prove that if a graph is planar, then so is any graph or multigraph obtained by contracting an edge.

\subsection{Subdivision}
There are two types of subdivision, an edge subdivision and graph subdivision. Edge subdivisions can best be though of as an operation, much like a contraction. We then can define a graph subdivision in terms of the edge version.

\subsubsection{Edge Subdivision}
An edge subdivision is where we take an an edge and add a vertex into the middle of it. If we have an edge $e$ that goes between vertex $a$ and $b$ then to subdivide it we add a vertex $c$, remove edge $e$ and then add an edge between $a$ and $c$ and between $b$ and $c$. This is demonstrated in the diagram below.
\begin{center}
	$
	\begin{array}{l}
		\begin{tikzpicture}[auto, node distance=3cm, every loop/.style={},
		thick,main node/.style={circle,draw,font=\sffamily\Large\bfseries}]

			\node[main node] (1) {$a$};
			\node[main node] (2) [right of=1] {$b$};

			\path[every node/.style={font=\sffamily\small}]
			(1)	edge node {$e$} (2);
		\end{tikzpicture}
	\end{array}
	$
	\scalebox{2}{$\to$}
	$
	\begin{array}{l}
		\begin{tikzpicture}[auto, node distance=3cm, every loop/.style={},
thick,main node/.style={circle,draw,font=\sffamily\Large\bfseries}]

		\node[main node] (1) {$a$};
		\node[main node] (3) [right of=1] {$c$};
		\node[main node] (2) [right of=3] {$b$};

		\path[every node/.style={font=\sffamily\small}]
		(1)	edge node {} (3)
		(2) edge node {} (3);
		\end{tikzpicture}
	\end{array}
	$
\end{center}

This can be thought of as a sort of inverse to contractions. To see this simply preform a contraction on either edge of the output graph above and you will return the the original graph.

\subsubsection{Graph Subdivision}
A graph $G$ is said to be a subdivision of $H$ if $G$ can be obtained from $H$ via some series of edge subdivisions. For example consider the two following graphs.
This can be thought of as just adding vertices along an edge. Intuitively this shouldn't actually change the shape of a graph. This not changing the shape of a graph is something that will be touched on later in lemma \ref{subdivision}.

@@

	\scalebox{.9}{$K_5$}
\end{center}
\begin{proof}
	$K_5$ is a graph\footnote{Recall that all graphs are multigraphs, but not all multigraphs are graphs.}, and as such an edge can not be a loop and two vertices can share at most one edge. This means that in any plane drawing of a graph, there must be at least 3 edges bordering every face.
	\begin{itemize}
		\item To only have one edge bordering a face would require that the edge be a loop (which we can't have in a graph).
		\item To only have only two edges bordering a face would require that some pair of vertices share more than one edge (which we can't have in a graph).
	\end{itemize}
	Additionally every edge boarders no more than two faces, so from this we know that in any graph $2e \ge 3f$ or $\frac23e\ge f$. Now $K_5$ has 10 edges and 5 vertices, so we know that $$\frac2310=\frac{20}3=6.\bar6\ge f$$ We know that $K_5$ is a connected graph (and thus also a connected multigraph) so if $K_5$ were planar then we would have $$2=v-e+f=5-10+f\le5-10+6.\bar6 = 1.\bar6$$ which is clearly false as $$2\not\le1.\bar6$$ Therefore $K_5$ must not be planar.
\end{proof}

@@

	\scalebox{.9}{$K_{3,3}$}
\end{center}
\begin{proof}
	$K_{3,3}$ is a bipartite graph, meaning that there is some way to break the graph into two collections of vertices, where there are no edges that stay within one of these collections. If we look at the drawing of $K_{3,3}$ above we see that the right and left works as these collections for us. There is no edge that goes between two vertices on the right or two vertices on the left. This means that if we can make a plane drawing of $K_{3,3}$ each face must have at least four edges on its boundary. We know from our proof of $K_5$'s nonplanarity that each face in a graph must have at least three edges on its boundary. If any face were to have three edges then the cycle bounding the face would have exactly three vertices as well. Each pair of vertices within these three vertices would share an edge and therefore there is no way to split them up into two groups such that neither group has an edge within it. Additionally every edge boarders no more than two faces, so from this we know that in any bipartite graph $2e \ge 4f$ or $e\ge 2f$.

@@

\begin{definition}[Drawing]
	A drawing is a subset, $D$, of some topological space $S$. For plane drawings this space will be $\mathbb R^2$ with it's standard topology.
\end{definition}

@@

\begin{lemma} \label{subdivision}
	If a graph $G$ is a subdivision of a nonplanar graph, then $G$ is nonplanar.
\end{lemma}

\begin{proof}
	Let $H$ be a nonplanar graph with edge, $e$, between vertices $a$ and $b$. Let $H'$ be the graph produced by preforming an edge subdivision on edge $e$. Let us call the new call the new vertex $v$ and the new edges $e_a$ and $e_b$ ($e_a$ is an edge from $a$ to $v$ and $e_b$ is an edge from $b$ to $v$).

	Suppose there exists a plane drawing, $D$, of $H'$. Then there are continuous injective functions $\hat e_a,\hat e_b\ccint01\to\mathbb R^2$ with $\hat e_a(1) = \hat e_b(0)$ whose range is contained in $D$. Additionally we know that $\hat e_a(x) \not= \hat e_b(y)$ for any $x,y\in\ooint 01$. Further we know that $a\not=b$ as otherwise $H$ would be a multigraph with $e$ being a loop in it. This means that $v(a) = \hat e_a(0) \not= \hat e_b(x)$ and $v(b) = \hat e_b(1)\not=\hat e_a(x)$ for all $x\in\ccint 01$. With all this we can construct an continuous injective function $\hat e:\ccint01\to \mathbb R^2$ such that $\hat e(0) = a$ and $\hat e(1) = b$ with 
	$$\hat e(x) = \begin{cases}
	\hat e_a(x*2) & x \le \frac12\\
	\hat e_b(x*2-1) & x > \frac 12
	\end{cases}$$
	As both $\hat e_a$ and $\hat e_b$ map into $D$, then so must $\hat e$ map into $D$. This contradicts our original statement that $H$ is nonplanar, and therefore $H'$ must not have a plane drawing.

	It follows by induction that if a graph is a subdivision of a nonplanar graph, that graph must be nonplanar.
\end{proof}

\begin{lemma} \label{component}
	If a graph $G$ is nonplanar than at least one of it's connected components is nonplanar.
\end{lemma}

\begin{proof}
	Suppose there exists a nonplanar graph $G$ for which all of it's components are planar. If each component is planar we may also make a drawing of it in any space homeomorphic to $\mathbb R^2$. Simply embed each component in a separate open epsilon ball about an integer point, with $\varepsilon \le \frac12$, and we find a contradiction with a planar embedding of $G$.
\end{proof}

\begin{lemma} \label{subdivisionCarryOn}
	If a graph $G$ has a subdivision of $H$ as a subgraph, then any subdivision of $G$ contains a subdivision of $H$ as a subgraph.
\end{lemma}
\begin{proof}
	Let $H$ be a graph. Let $H'$ be a subdivision of $H$. Let $G$ be a graph that contains $H'$ as a subgraph. Let $G'$ be a subdivision of $G$. There is some sequence of edge subdivisions $h_1,h_2,\ldots,h_\ell$ that produce $H'$ from $H$ as is there some sequence of edge subdivisions $g_1,g_2,\ldots,g_m$ that produce $G'$ from $G$. Let $g_{i_1},g_{i_2},\ldots,g_{i_n}$ be the subsequence of all the edge subdivisions that happen within $H'$. We then can produce $H''$ by taking $H$ and doing the sequence of edge subdivisions $h_1,h_2,\ldots,h_\ell,g_{i_1},g_{i_2},\ldots,g_{i_n}$ so $H''$ is a subdivision of $H$.
\end{proof}

@@

\begin{definition}[Jordan curve]
	A Jordan curve is a curve, $f:\ccint01\to \mathbb R^2$, with the following properties:
	\begin{itemize}
		\item $f$ is continuous
		\item $f$ is injective on the interval $\coint 01$
		\item $f(0) = f(1)$
	\end{itemize}
\end{definition}
\begin{lemma}
	For any plane drawing of a graph, $G$, containing cycle $C$, $C$ is a Jordan curve in the drawing.
\end{lemma}
\begin{proof}
%TODO ask Shaw what he thinks
	Consider a plane drawing of a graph $G$ with cycle $C$. We may write out $C$ as $v_1e_1v_2e_2\ldots v_ke_kv_1$. We know that each of these $e_i$s have a corresponding curve $\hat e_i$ that starts at the end of the previous ($\hat e_{i-1}$) curve and ends at the beginning of the next ($\hat e_{i+1}$) curve. This means we can construct a new continuous curve simply by stringing each of these together and then resize the domain. What we end with will then be a curve $f:\ccint 01\to\mathbb R^2$ that is continuous. We will also have $f(0) = f(1)$ as $v_1$ as $e_1$ starts at $v_1$ and $e_k$ ends at $v_1$. We will also have $f$ injective on $\coint{0}{1}$ as we know that edges do not overlap each other (this is formalized in the fourth bullet point of our definition for a drawing). This then means that our cycle creates a Jordan curve.
\end{proof}

@@

\begin{lemma} \label{outside}
	For any face, $F$, in a plane drawing, $D$, of a graph $G$, there is a corresponding drawing plane drawing of $G$, $D'$, where all edges and vertices that boarder $F$ in $D$, now boarder the outer face.
\end{lemma}
\begin{proof}
	Let $G$ be a graph with face $F$ in plane drawing $D$. Now let $\inv f$ be a stereographic projection from the punctured sphere onto the plane. Therefore $f(D)$ is a drawing of $G$ on the punctured sphere. If $f(D)$ is a drawing of $G$ on the punctured sphere then it must also be a drawing of $G$ on the sphere. Now we pick a point on the interior of $F$ and puncture the sphere there rather than where our original puncture was.\footnote{There is a problem in this proof. How do we know that $F$ has an interior? Simply put if we have a finite number of vertices and never use any space filling curves for edges we will never fill space. We know that we are never forced to use a space filling curve as for any space filling curve $c$ we may create a new curve $c'$ that connects the same end points, stays within $c$'s image, and is not space filling. %TODO ask Shaw if maybe turn into lemma and prove
	} We may now use a sterographic projection, $g$ to map $f(D)$ from the new punctured sphere back to $\mathbb R^2$ creating $D' = g(f(D))$. We know that in $D'$ the face that corresponds to $F$ is the outer face as it contained the punctured point in $\inv g(D')$. %TODO Ask shaw if I should prove that the puctured point in the sphere corresponds to a limit point
\end{proof}

@@

\begin{definition}[$k$ connected]
	A graph, $G$, is $k$ connected if it's minimal cut set is of size $k$. The complete graph on $n$ vertices is defined to be $n-1$ connected.
\end{definition}

@@

\begin{definition}[mutually disjoint paths]
	A set of paths, $S$, is mutually disjoint if no two paths share a vertex. A path $p\in S$ is said to be mutually disjoint from the the other paths in $S$ if $S$ is mutually disjoint.
\end{definition}

\begin{definition}[mutually internally disjoint paths]
	A set of paths, $S$, is mutually internally disjoint if no to paths share an internal vertex. That is no two paths, $p_1\not=p_2\in S$ share a vertex other than their first or last vertex. A path $p\in S$ is said to be mutually internally disjoint from the other paths in $S$ if $S$ is mutually internally disjoint.
\end{definition}

\begin{theorem}[Menger's theorem]
	A non-empty graph is $n$ connected if and only if among all pairs of vertices $a\not=b$ in $G$ the least number of mutually internally disjoint paths from $a$ to $b$ is $n$.
\end{theorem}
%TODO cite Wikipeida menger's theorem
\begin{proof}
	This is trivially true on complete graphs, so for the rest of this proof we will only be considering graphs that are not complete.

>>>>

%\subsection{Neighborhood}
%Given some graph $G$ with vertex $v$. The neighborhood of $v$ within $G$ is the set of all vertices that share an edge with $v$, not including $v$.\footnote{If $G$ is a multigraph the definition stays the same, however $v$ maybe its own neighbor if it has an edge with itself or a loop} Given a subgraph $H$ of $G$, the neighborhood of $H$ within $G$ is a new subgraph of $G$ that includes $H$ as well as any vertex in $G$ that is adjacent to a vertex in $H$ and the edge that they share.
%
%Lets look at some examples
%\begin{center}
%	\begin{tikzpicture}[auto, node distance=3cm, every loop/.style={},
%	thick,main node/.style={circle,draw,font=\sffamily\Large\bfseries}]
%	
%		\node[main node] (0) {\textcolor{blue}{$a$}};
%		\node[main node] (1) [above right of=0] {\textcolor{red}{$b$}};
%		\node[main node] (2) [right of=1] {$c$};
%		\node[main node] (3) [below right of=2] {$d$};
%		\node[main node] (4) [below of=3] {\textcolor{red}{$e$}};
%		\node[main node] (5) [below left of=4] {$f$};
%		\node[main node] (6) [left of=5] {$g$};
%		\node[main node] (7) [above left of=6] {$h$};
%		
%		\path[every node/.style={font=\sffamily\small}]
%		(0)
%		edge node {} (2)
%		edge node {} (3)
%		edge node {} (4)
%		(1)
%		edge node {} (3)
%		edge node {} (4)
%		edge node {} (5)
%		(2)
%		edge node {} (4)
%		edge node {} (5)
%		(3)
%		edge node {} (7)
%		(4)
%		edge node {} (6)
%		edge node {} (7)
%		(5)
%		edge node {} (6)
%		edge node {} (7)
%		(6)
%		edge node {} (7);
%	\end{tikzpicture}
%\end{center}
%
%First let us consider the neighborhood of just the vertex \textcolor{blue}{$a$}, this would then be all the vertices that it is adjacent to so we get $$\set{c, d, \textcolor{red}{e}}$$ We could also let $A$ be a subgraph containing only the vertex $\textcolor{blue}{a}$ and so its neighborhood within $G$ will be the graph below.
%\begin{center}
%	\begin{tikzpicture}[auto, node distance=3cm, every loop/.style={},
%	thick,main node/.style={circle,draw,font=\sffamily\Large\bfseries}]
%	
%		\node[main node] (0) {\textcolor{blue}{$a$}};
%		\node[main node] (2) [above right of=0] {$c$};
%		\node[main node] (3) [below right of=2] {$d$};
%		\node[main node] (4) [below of=3] {\textcolor{red}{$e$}};
%		
%		\path[every node/.style={font=\sffamily\small}]
%		(0)
%		edge node {} (2)
%		edge node {} (3)
%		edge node {} (4)
%		(2)
%		edge node {} (4);
%	\end{tikzpicture}
%\end{center}
%The difference is that this is actually a subgraph of $G$, whereas the previous is just vertices.
%
%For another example let $H$ be the subgraph with only the \textcolor{red}{$b$}, \textcolor{red}{$e$} and the edge they share. The neighborhood of $H$ within $G$ within $G$ would then be all of $G$ as every vertex is either a neighbor of $b$ or $e$.
%
\subsection{Contraction}

\subsubsection{Edge contraction}
This is not a property, but rather an operation or action that we perform on a graph. Let us consider the following graph.

@@

We wish to perform a contraction on edge $R$. To be clear all graph contractions are on edges. So we will make a new vertex, $r\cdot s$ which has all the edges of $r$ and all the edges of $s$ except the edge we are contracting across. In this case the edge we are contracting across is $R$ and thus we get the following multigraph.

@@

Often we may prefer to have the end result be a graph rather than a multigraph, which brings us to our next kind of contraction

\subsubsection{Simple edge contraction}

We can take any multigraph produced by an edge contraction and for any edge pairs that go between the same vertices, simply replace it with a single edge. This will reduce the multigraph back to a graph, and from the previous example the result of a simple edge contraction on $R$ will give the below graph.

	\path[every node/.style={font=\sffamily\small}]
	(1)	edge node {$a$} (3)
	(2) edge node {$b$} (1)
	(4) edge node {$c$} (1);
	\end{tikzpicture}
\end{center}

%\subsection{Subdivision}
%There are two types of subdivision, an edge subdivision and graph subdivision. Edge subdivisions can best be though of as an operation, much like a contraction. We then can define a graph subdivision in terms of the edge version.

%\subsubsection{Edge Subdivision}
%An edge subdivision is where we take an an edge and add a vertex into the middle of it. If we have an edge $e$ that goes between vertex $a$ and $b$ then to subdivide it we add a vertex $c$, remove edge $e$ and then add an edge between $a$ and $c$ and between $b$ and $c$. This is demonstrated in the diagram below.
%\begin{center}
%	$
%	\begin{array}{l}
%		\begin{tikzpicture}[auto, node distance=3cm, every loop/.style={},
%		thick,main node/.style={circle,draw,font=\sffamily\Large\bfseries}]
%		
%			\node[main node] (1) {$a$};
%			\node[main node] (2) [right of=1] {$b$};
%			
%			\path[every node/.style={font=\sffamily\small}]
%			(1)	edge node {$e$} (2);
%		\end{tikzpicture}
%	\end{array}
%	$
%	\scalebox{2}{$\to$}
%	$
%	\begin{array}{l}
%		\begin{tikzpicture}[auto, node distance=3cm, every loop/.style={},
%thick,main node/.style={circle,draw,font=\sffamily\Large\bfseries}]
%
%		\node[main node] (1) {$a$};
%		\node[main node] (3) [right of=1] {$c$};
%		\node[main node] (2) [right of=3] {$b$};
%		
%		\path[every node/.style={font=\sffamily\small}]
%		(1)	edge node {} (3)
%		(2) edge node {} (3);
%		\end{tikzpicture}
%	\end{array}
%	$
%\end{center}
%
%This can be thought of as a sort of inverse to contractions. To see this simply perform a contraction on either edge of the output graph above and you will return the the original graph.
%
%
%\subsubsection{Graph Subdivision}
%A graph $G$ is said to be a subdivision of $H$ if $G$ can be obtained from $H$ via some series of edge subdivisions. For example consider the two following graphs.
%This can be thought of as just adding vertices along an edge. Intuitively this shouldn't actually change the shape of a graph. This not changing the shape of a graph is something that will be touched on later in lemma \ref{subdivision}.

@@

	\scalebox{.9}{$K_5$}
\end{center}
\begin{proof}
	$K_5$ is a graph\footnote{Recall that all graphs are multigraphs, but not all multigraphs are graphs.}, and as such an edge can not be a loop and two vertices can share at most one edge. This means that in any plane drawing of a graph, there must be at least 3 edges bordering every face.
	\begin{itemize}
		\item To only have one edge bordering a face would require that the edge be a loop (which we can't have in a graph).
		\item To only have only two edges bordering a face would require that some pair of vertices share more than one edge (which we can't have in a graph).
	\end{itemize}
	Additionally every edge borders no more than two faces, so from this we know that in any graph $2e \ge 3f$ or $\frac23e\ge f$. Now $K_5$ has 10 edges and 5 vertices, so we know that $$\frac2310=\frac{20}3=6.\bar6\ge f$$ We know that $K_5$ is a connected graph (and thus also a connected multigraph) so if $K_5$ were planar then we would have $$2=v-e+f=5-10+f\le5-10+6.\bar6 = 1.\bar6$$ which is clearly false as $$2\not\le1.\bar6$$ Therefore $K_5$ must not be planar.
\end{proof}

@@

	\scalebox{.9}{$K_{3,3}$}
\end{center}
\begin{proof}
	$K_{3,3}$ is a bipartite graph, meaning that there is some way to break the graph into two collections of vertices, where there are no edges that stay within one of these collections. If we look at the drawing of $K_{3,3}$ above we see that the right and left works as these collections for us. There is no edge that goes between two vertices on the right or two vertices on the left. This means that if we can make a plane drawing of $K_{3,3}$ each face must have at least four edges on its boundary. We know from our proof of $K_5$'s nonplanarity that each face in a graph must have at least three edges on its boundary. If any face were to have three edges then the cycle bounding the face would have exactly three vertices as well. Each pair of vertices within these three vertices would share an edge and therefore there is no way to split them up into two groups such that neither group has an edge within it. Additionally every edge borders no more than two faces, so from this we know that in any bipartite graph $2e \ge 4f$ or $e\ge 2f$.

@@

\begin{definition}[Drawing]
	A drawing is a subset, $D$, of some topological space $S$. For plane drawings this space will be $\mathbb R^2$ with its standard topology.
\end{definition}

@@

\begin{definition}[contactable]
	A graph $G$ is said to be contactable to a graph $H$ if there is some series of simple edge contractions that when performed on $G$ produce $H$.
\end{definition}

\begin{lemma}
	If a graph $G$ with edge $e$ is planar then so is the graph produced by performing a simple edge contraction on $e$.
\end{lemma}

\begin{proof}
	Let $G$ be a planar graph with edge $e$ between $x$ and $y$. et $G'$ be the graph produced by an edge contraction on $e$. We are going to prove this using induction on the number of edges of $y$.

	Statement: $G'$ is planar and there is a plane drawing of $G'$ where all vertices, $v$, that border a face that $y$ borders in $G$ also a face that $x\cdot y$ borders in $G'$ ($x\cdot y$ is the vertex produced by contracting across $e$).

	Base case: Suppose $y$ only has edge $e$. Clearly $G'\subset G$ so by lemma \ref{subgraph} $G'$ must be planar. Additionally as $y$ has only one edge it can not possibly lie upon any cycles in $G$, this means that $y$ can only border at most 1 face, and as it has an edge to $x$ it must share this face with $x$. We may draw $G'$ simply be removing $e$ and $y$ from any drawing of $G$ and relabeling $x$ as $x\cdot y$. Clearly then the statement above must be true when $y$ has only one edge.

	Inductive step: Let $z\not=x$ be a vertex sharing edge $\bar e$ with $y$. Assume that the statement is true in $G\setminus\set{\bar e}$. Now perform a contraction on $e$ in $G\setminus\set{\bar e}$ to produce $H$. $x\cdot y$ in $H$ must border a face $\bar f$ that $z$ borders so we may add in edge $\bar e$ to $H$ in face $\bar f$ and produce a plane drawing of $G'$. Further let $v$ be a vertex bordering a face $f$ in $G$ that $y$ also borders. We then know that there is a face $f'$ in $H$ that borders both $v$ and $x \cdot y$. If $f'\not=\bar f$ then clearly $x\cdot y$ and $v$ both boarder $f'$ in $G'$ as well as $H$. If $f'=\bar f$ then $f'$ is split into two parts by $\bar e$ in $G'$, both of which are bordered by $x\cdot y$. Therefore our statement must be true for $G$.

	Now by induction we have proven the statement is true on all graphs and therefore our lemma is proved.
\end{proof}

\begin{corallary} \label{contractible}
	If a graph $G$ is planar, and if $G$ is contactable to a graph $H$, then $H$ must also be planar.
\end{corallary}

\begin{proof}
	Let $c_1,c_2,\ldots, c_n$ be the sequence of contractions on $G$ that produce $H$. Let $H_i$ be the graph produced after the first $i$ contractions on $G$, so $H_i = G$ and $H_n = H$. If $G$ is planar then clearly $H_0$ is planar. If $H_i$ is planar, then by the above lemma clearly $H_{i+1}$ is planar. Now by induction, if $G$ is planar $H$ must be planar.
\end{proof}

\begin{lemma} \label{contractibleCarryOn}
	If a graph $G$ has a subgraph that is contactable to $H$ as a subgraph, then graph $G'$ that is contactable to $G$ also contains a subgraph that is contactable to $H$.
\end{lemma}

\begin{proof}
		Let $H$ be a graph. Let $H'$ be contactable to $H$. Let $G$ be a graph that contains $H'$ as a subgraph. Let $G'$ be a graph that is contactable to $G$. There is some sequence of simple edge contractions $h_1,h_2,\ldots,h_\ell$ that produce $H$ from $H'$ as is there some sequence of simple edge contractions $g_1,g_2,\ldots,g_m$ that produce $G$ from $G'$.
		If we reverse these sequences we get sequences of inverse simple edge contractions that take produce $H'$ from $H$ and $G'$ from $G$. Each inverse simple edge contraction will be done on a single vertex (and turn it into two vertices that share an edge). Let $f_i = h_{\ell-i+1}$ and $e_i = g_{i-m+1}$ so that they are the reversed sequences of $g$ and $h$. Take the subsequence, $e_{i_1},e_{i_2},\ldots,e_{i_n}$ that consists of all the inverse simple edge contractions that act on a vertex within the subgraph $H'$ of $G$. Now we can produce $H''$ as the graph produced by performing the following inverse simple edge contractions on $H$
		$$ f_1,f_2,\ldots,f_\ell,e_{i_1},e_{i_2},\ldots,e_{i_n} $$
		Clearly now $H''$ is contactable to $H$ and $H''$ must be a subgraph of $G'$.
\end{proof}

%%TODO remove?
%\begin{lemma} \label{subdivision}
%	If a graph $G$ is a subdivision of a nonplanar graph, then $G$ is nonplanar.
%\end{lemma}
%
%\begin{proof}
%	Let $H$ be a nonplanar graph with edge, $e$, between vertices $a$ and $b$. Let $H'$ be the graph produced by performing an edge subdivision on edge $e$. Let us call the new call the new vertex $v$ and the new edges $e_a$ and $e_b$ ($e_a$ is an edge from $a$ to $v$ and $e_b$ is an edge from $b$ to $v$).
%	
%	Suppose there exists a plane drawing, $D$, of $H'$. Then there are continuous injective functions $\hat e_a,\hat e_b\ccint01\to\mathbb R^2$ with $\hat e_a(1) = \hat e_b(0)$ whose range is contained in $D$. Additionally we know that $\hat e_a(x) \not= \hat e_b(y)$ for any $x,y\in\ooint 01$. Further we know that $a\not=b$ as otherwise $H$ would be a multigraph with $e$ being a loop in it. This means that $v(a) = \hat e_a(0) \not= \hat e_b(x)$ and $v(b) = \hat e_b(1)\not=\hat e_a(x)$ for all $x\in\ccint 01$. With all this we can construct an continuous injective function $\hat e:\ccint01\to \mathbb R^2$ such that $\hat e(0) = a$ and $\hat e(1) = b$ with 
%	$$\hat e(x) = \begin{cases}
%	\hat e_a(x*2) & x \le \frac12\\
%	\hat e_b(x*2-1) & x > \frac 12
%	\end{cases}$$
%	As both $\hat e_a$ and $\hat e_b$ map into $D$, then so must $\hat e$ map into $D$. This contradicts our original statement that $H$ is nonplanar, and therefore $H'$ must not have a plane drawing.
%	
%	It follows by induction that if a graph is a subdivision of a nonplanar graph, that graph must be nonplanar.
%\end{proof}

\begin{lemma} \label{component}
	If a graph $G$ is nonplanar than at least one of its connected components is nonplanar.
\end{lemma}

\begin{proof}
	Suppose there exists a nonplanar graph $G$ for which all of its components are planar. If each component is planar we may also make a drawing of it in any space homeomorphic to $\mathbb R^2$. Simply embed each component in a separate open epsilon ball about an integer point, with $\varepsilon \le \frac12$, and we find a contradiction with a planar embedding of $G$.
\end{proof}

%%TODO remove?
%\begin{lemma} \label{subdivisionCarryOn}
%	If a graph $G$ has a subdivision of $H$ as a subgraph, then any subdivision of $G$ contains a subdivision of $H$ as a subgraph.
%\end{lemma}
%\begin{proof}
%	Let $H$ be a graph. Let $H'$ be a subdivision of $H$. Let $G$ be a graph that contains $H'$ as a subgraph. Let $G'$ be a subdivision of $G$. There is some sequence of edge subdivisions $h_1,h_2,\ldots,h_\ell$ that produce $H'$ from $H$ as is there some sequence of edge subdivisions $g_1,g_2,\ldots,g_m$ that produce $G'$ from $G$. Let $g_{i_1},g_{i_2},\ldots,g_{i_n}$ be the subsequence of all the edge subdivisions that happen within $H'$. We then can produce $H''$ by taking $H$ and doing the sequence of edge subdivisions $h_1,h_2,\ldots,h_\ell,g_{i_1},g_{i_2},\ldots,g_{i_n}$ so $H''$ is a subdivision of $H$.
%\end{proof}

@@

\begin{definition}[Jordan curve]
	A Jordan curve is a curve, $f:\ccint01\to \mathbb R^2$, with the following properties:
	\begin{itemize}
		\item $f$ is continuous
		\item $f$ is injective on the interval $\coint 01$
		\item $f(0) = f(1)$
	\end{itemize}
\end{definition}
\begin{lemma}\label{JordansLemma}
	For any plane drawing of a graph, $G$, containing cycle $C$, $C$ is a Jordan curve in the drawing.
\end{lemma}
\begin{proof}
%TODO ask Shaw what he thinks
	Consider a plane drawing of a graph $G$ with cycle $C$. We may write out $C$ as $v_1e_1v_2e_2\ldots v_ke_kv_1$. We know that each of these $e_i$s have a corresponding curve $\hat e_i$ that starts at the end of the previous ($\hat e_{i-1}$) curve and ends at the beginning of the next ($\hat e_{i+1}$) curve. This means we can construct a new continuous curve simply by stringing each of these together and then resize the domain. What we end with will then be a curve $f:\ccint 01\to\mathbb R^2$ that is continuous. We will also have $f(0) = f(1)$ as $v_1$ as $e_1$ starts at $v_1$ and $e_k$ ends at $v_1$. We will also have $f$ injective on $\coint{0}{1}$ as we know that edges do not overlap each other (this is formalized in the fourth bullet point of our definition for a drawing). This then means that our cycle creates a Jordan curve.
\end{proof}

@@

\begin{lemma} \label{outside}
	For any face, $F$, in a plane drawing, $D$, of a graph $G$, there is a corresponding drawing plane drawing of $G$, $D'$, where all edges and vertices that border $F$ in $D$, now border the outer face.
\end{lemma}
\begin{proof}
	Let $G$ be a graph with face $F$ in plane drawing $D$. Now let $\inv f$ be a stereographic projection from the punctured sphere onto the plane. Therefore $f(D)$ is a drawing of $G$ on the punctured sphere. If $f(D)$ is a drawing of $G$ on the punctured sphere then it must also be a drawing of $G$ on the sphere. Now we pick a point on the interior of $F$ and puncture the sphere there rather than where our original puncture was.\footnote{There is a problem in this proof. How do we know that $F$ has an interior? Simply put if we have a finite number of vertices and never use any space filling curves for edges we will never fill space. We know that we are never forced to use a space filling curve as for any space filling curve $c$ we may create a new curve $c'$ that connects the same end points, stays within $c$'s image, and is not space filling. %TODO ask Shaw if maybe turn into lemma and prove
	} We may now use a sterographic projection, $g$ to map $f(D)$ from the new punctured sphere back to $\mathbb R^2$ creating $D' = g(f(D))$. We know that in $D'$ the face that corresponds to $F$ is the outer face as it contained the punctured point in $\inv g(D')$. %TODO Ask shaw if I should prove that the puctured point in the sphere corresponds to a limit point
\end{proof}

@@

\begin{definition}[$k$ connected]
	A graph, $G$, is $k$ connected if its minimal cut set is of size $k$. The complete graph on $n$ vertices is defined to be $n-1$ connected.
\end{definition}

@@

\begin{definition}[disjoint paths]
	A set of paths, $S$, is disjoint if no two paths share a vertex. A path $p\in S$ is said to be disjoint from the the other paths in $S$ if $S$ is disjoint.
\end{definition}

\begin{definition}[internally disjoint paths]
	A set of paths, $S$, is internally disjoint if no to paths share an internal vertex. That is no two paths, $p_1\not=p_2\in S$ share a vertex other than their first or last vertex. A path $p\in S$ is said to be internally disjoint from the other paths in $S$ if $S$ is internally disjoint.
\end{definition}

\begin{theorem}[Menger's theorem]\label{menger}
	A non-empty graph is $n$ connected if and only if among all pairs of vertices $a\not=b$ in $G$ the least number of internally disjoint paths from $a$ to $b$ is $n$.
\end{theorem}
%TODO cite Wikipeida menger's theorem
\begin{proof}
	This is trivially true on complete graphs, so for the rest of this proof we will only be considering graphs that are not complete.

	Let $A$ and $B$ be vertex sets in a finite graph $G$. Define an $AB$ path as any path that starts in $A$ and has no other vertices in $A$, and ends in $B$ and has no other vertices in $B$. Notice that if $v\in A\inter B$ then the path $v$ is an $AB$ path. Define an $AB$ separator as any vertex set, $S$, for which all $AB$ paths contain at least one member of $S$. Define an $AB$ connector a set of disjoint $AB$ paths.
